The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-11-20 Manuscript_draft/Draft.Rmd
Other changes in the draft

The scaling parameter $A$ was still needed to explain some of the variance in respiration measured at the level of single experimental plots even after normalization by SOC stocks. This is likely because of the high variability in SOC content of soils

## Role of temperature and moisture in decomposition 
The process-based model showed that processes related to soil temperature and moisture were able to explained more than 50% of variance in soil respiration. MORE REFS RELATED TO MODELS (Nissan et al 2023,etc.).

@@

### Relationships between organic matter charachteristics, inputs of C to the soil, and temperature sensitivity

@@

Our results also confirmed the hypothesis 2, since addition of slash decreased the activation energy of SOM decomposition. We argue that this result is probably related to a reduction in chemical recalcitrance, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999). Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens (2006) and Sierra (2012) that the higher the activation energy the higher is the relative temperature sensitivity. We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites. This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014). Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases in activation energies in the plots where removal of slash had an effect. Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022). In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant. Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022). This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance (Levegre et al 2014).

>>>>

The scaling parameter $A$ was still needed to explain some of the variance in respiration measured at the level of single experimental plots even after normalization by SOC stocks. This is likely because of the high variability in SOC content of soils, which, particularly in forest soils, can be affected by local processes at a very small spatial scale. For example small depressions and local micromorphology, root concentration, localized events such as branch or tree falls, and so on, can substantially affect SOC at a metric or submetric scale. This could easily cause discrepancies between the SOC stocks at the site of each respiration measuring collar and the SOC stocks measured by pooling at the level of each experimental plot. Since respiration depends first of all on the amoutn of substrate, local stocks can heavily affect it. By allowing the $A$ parameter to vary at the level of the experimental plot we could allow for the needed freedom for our parametric models to account for this effect (at the cost of an increased variance of the posterior, but reducing eventual biases).

## Role of temperature and moisture in decomposition 
The process-based model showed that processes related to soil temperature and moisture were able to explain more than 50% of variance in soil respiration. MORE REFS RELATED TO MODELS (Nissan et al 2023,etc.).

@@

@@

Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly when in a clearc cut (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots that start to decompose). We argue that this result is probably related to a reduction in chemical recalcitrance due to the increased amount of fresh C inputs, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool [@Bosatta.1999], confirmed by other direct measurements [@Lefevre.2014]. A more recalcitrant substrate likely implies more biochemical steps for its decomposition, each step subjected to the same thermodynamic constraints. More steps in the overall oxydative reaction leading also implies a higher chance of one being more limited than the others due to the enzyme characteristics.

We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites. This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014). Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases in activation energies in the plots where removal of slash had an effect. Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davidson and Jansen 2022). In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant. Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022). This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance.
